---
title: "HAZ Simulation — Improved (K=1000, Worst-Case & Residual Variance Scenarios)"
author: "Mthobisi Zondi"
date: "2026-03-10"
output: html_document
---

<!--
  DESIGN NOTES
  ============
  Outcome  : Height-for-Age Z-score (HAZ)
  Cohort   : HIV-Uninfected (HUU), N_TARGET = 736
  Reference country: Malawi (all country fixed effects are
                     deviations from Malawi's intercept)

  True fixed-effect parameters
  ─────────────────────────────
    β₀  Intercept              −0.50
    β₁  Linear age  (g1)       −0.015
    β₂  Log age     (g2)        0.30
    β₃  Inverse age (g3)       −0.10
    β_SA South Africa vs Malawi  0.45
    β_UG Uganda vs Malawi        0.20
    β_ZW Zimbabwe vs Malawi      0.15

  Sample-size scenarios (multiples of real N = 736)
  ──────────────────────────────────────────────────
    N =  368  (½ × real — under-recruitment / high dropout)
    N =  736  (1 × real — Study Anchor)
    N = 1472  (2 × real — pooled / expanded replication)
    N = 3680  (5 × real — registry scale)

  Random-effects scenario grid (16 scenarios, 4 groups)
  ──────────────────────────────────────────────────────
    Group A: Variance magnitude  (σ₀ and σ₁ scale together)
    Group B: Slope variability   (σ₁ varies; σ₀ fixed at base)
    Group C: Intercept–slope ρ   (ρ varies; variances fixed at base)
    Group D: Worst-case combined (extreme variance + high correlation)
             NEW: stress-tests LMM under the most challenging conditions

  Residual variance scenarios (Group E) — NEW
  ─────────────────────────────────────────────
    σ_ε varies from 0.40 to 1.60 to assess sensitivity to
    measurement noise (e.g., anthropometry error, secular variation).

  Reporting metrics — EXPANDED
  ──────────────────────────────
    • Bias           : E[β̂] − β_true
    • RMSE           : √E[(β̂ − β_true)²]
    • Coverage       : % of 95% CIs containing β_true         ← NEW
    • Convergence    : % of reps where LMM converged           ← NEW
    • Type I error   : % false positives when β = 0 (null)    ← NEW
    • Power          : % of reps correctly rejecting H₀        ← NEW
                       Reported as a function of N and effect
                       size (country contrasts and age terms).
                       Power curve fitted via logistic regression
                       on the binary reject/not-reject outcome.

  OPTIMISATION NOTES
  ==================
  • R_REPS_MAIN  = 1000  (was 200) — full precision for bias/RMSE/coverage
  • R_REPS_CROSS =  200  (was  50) — crossed grid
  • Cache directory: sim_cache_HAZ/ — each scenario saved as .rds
  • To restore fast-run mode: R_REPS_MAIN=200, R_REPS_CROSS=50
-->

---

# 0. Packages & Global Settings

```{r chunk-0-packages, message=FALSE, warning=FALSE}
packages <- c(
  "lme4", "lmerTest", "MASS", "dplyr", "tidyr",
  "ggplot2", "knitr", "broom.mixed", "pbapply", "purrr", "data.table"
)

installed <- rownames(installed.packages())
for (pkg in packages) {
  if (!pkg %in% installed) install.packages(pkg, quiet = TRUE)
}

library(lme4);      library(lmerTest); library(MASS)
library(dplyr);     library(tidyr);    library(ggplot2)
library(knitr);     library(broom.mixed)
library(pbapply);   library(purrr);    library(data.table)

set.seed(20260220)

# ── Study anchor ──────────────────────────────────────────────────────────────
N_TARGET <- 736   # actual number of individuals in the real HUU dataset

# ── Replication counts ────────────────────────────────────────────────────────
# CHANGED: R_REPS_MAIN raised from 200 → 1000 for stable coverage estimates.
# Coverage of a 95% CI requires ~1000 reps to estimate with SE < 0.7 pp.
# Formula: SE(coverage) = sqrt(0.95 * 0.05 / R) → SE < 0.007 when R = 1000.
R_REPS_MAIN  <- 1000   # ← WAS 200; full precision for bias/RMSE/coverage
R_REPS_CROSS <-  200   # ← WAS  50; crossed (N × RE) grid

# ── HAZ-specific cache directory ─────────────────────────────────────────────
CACHE_DIR <- "sim_cache_HAZ"
if (!dir.exists(CACHE_DIR)) dir.create(CACHE_DIR)

cache_save <- function(obj, name) {
  saveRDS(obj, file.path(CACHE_DIR, paste0(name, ".rds")))
  invisible(obj)
}
cache_load <- function(name) {
  path <- file.path(CACHE_DIR, paste0(name, ".rds"))
  if (file.exists(path)) readRDS(path) else NULL
}
```

---

# 1. Data-Generating Function

```{r chunk-1-data-gen}
# ── HAZ data generator ────────────────────────────────────────────────────────
# NEW parameter: sigma_eps (residual SD) — previously hard-coded to 0.80.
# Exposing it enables Group E (residual variance) scenarios.
# ─────────────────────────────────────────────────────────────────────────────
generate_HAZ_data <- function(
    N         = 500,
    dropout   = 0.20,
    sigma_b0  = 0.6,
    sigma_b1  = 0.02,
    rho_b01   = 0.2,
    sigma_eps = 0.80   # ← NEW: residual SD (was hard-coded)
) {

  # 1. Country assignment
  country <- sample(
    c("Uganda", "Malawi", "Zimbabwe", "South Africa"),
    size    = N, replace = TRUE,
    prob    = c(117, 234, 172, 75) / 598
  )

  # 2. Maternal age
  age_mean <- c(Uganda = 31.44, Malawi = 31.57,
                Zimbabwe = 32.75, "South Africa" = 32.51)
  age_sd   <- c(Uganda =  5.43, Malawi =  5.13,
                Zimbabwe =  5.44, "South Africa" =  5.46)
  maternal_age <- rnorm(N, age_mean[country], age_sd[country])

  # 3. Education
  edu_levels <- c("no_school", "primary_partial", "primary_complete",
                  "secondary_partial", "secondary_complete", "college")
  edu_probs  <- list(
    Uganda        = c(0.0342, 0.0940, 0.2821, 0.4359, 0.0855, 0.0684),
    Malawi        = c(0.0855, 0.1410, 0.3333, 0.2991, 0.1154, 0.0256),
    Zimbabwe      = c(0.0000, 0.0698, 0.0465, 0.3314, 0.5523, 0.0000),
    "South Africa"= c(0.0000, 0.0133, 0.0400, 0.5067, 0.2933, 0.1467)
  )
  edu_cat  <- vapply(country,
                     function(c) sample(edu_levels, 1, prob = edu_probs[[c]]),
                     character(1))
  edu_high <- as.integer(edu_cat %in% c("secondary_complete", "college"))

  # 4. Employment
  work_levels <- c("not_employed", "self_employed", "formal")
  work_probs  <- list(
    Uganda        = c(0.3846, 0.3932, 0.2222),
    Malawi        = c(0.5556, 0.3248, 0.1197),
    Zimbabwe      = c(0.4942, 0.4070, 0.0988),
    "South Africa"= c(0.6800, 0.0400, 0.2800)
  )
  work <- vapply(country,
                 function(c) sample(work_levels, 1, prob = work_probs[[c]]),
                 character(1))

  # 5. Marital status
  marital_levels <- c("no_partner", "primary_partner", "married",
                      "separated", "divorced", "widowed")
  marital_probs  <- list(
    Uganda        = c(0.0598, 0.3590, 0.5128, 0.0684, 0.0000, 0.0000),
    Malawi        = c(0.0085, 0.0128, 0.8333, 0.0214, 0.0470, 0.0769),
    Zimbabwe      = c(0.0174, 0.0465, 0.8547, 0.0349, 0.0116, 0.0349),
    "South Africa"= c(0.1333, 0.7333, 0.1200, 0.0133, 0.0000, 0.0000)
  )
  marital <- vapply(country,
                    function(c) sample(marital_levels, 1, prob = marital_probs[[c]]),
                    character(1))

  # 6. Infrastructure
  elec_prob <- c(Uganda = 0.7949, Malawi = 0.5385,
                 Zimbabwe = 0.5930, "South Africa" = 0.9867)
  tap_prob  <- c(Uganda = 0.3761, Malawi = 0.6154,
                 Zimbabwe = 0.5698, "South Africa" = 0.9890)
  electricity <- rbinom(N, 1, elec_prob[country])
  tapwater    <- rbinom(N, 1,  tap_prob[country])

  # 7. Toilet
  toilet_levels <- c("none", "other", "pit_latrine",
                     "private_flush", "shared_flush")
  toilet_probs  <- list(
    Uganda        = c(0.0256, 0.0000, 0.8793, 0.0259, 0.0862),
    Malawi        = c(0.0052, 0.0000, 0.9430, 0.0311, 0.0207),
    Zimbabwe      = c(0.0116, 0.0291, 0.1977, 0.2267, 0.5349),
    "South Africa"= c(0.0000, 0.0000, 0.1200, 0.6400, 0.2400)
  )
  toilet <- vapply(country,
                   function(c) sample(toilet_levels, 1, prob = toilet_probs[[c]]),
                   character(1))

  # 8. Travel time to clinic
  travel_levels <- c("<30min", "30_60min", ">1hr")
  travel_probs  <- list(
    Uganda        = c(0.0342, 0.3675, 0.5983),
    Malawi        = c(0.0684, 0.5256, 0.4060),
    Zimbabwe      = c(0.4593, 0.3605, 0.1802),
    "South Africa"= c(0.3733, 0.4267, 0.2000)
  )
  travel <- vapply(country,
                   function(c) sample(travel_levels, 1, prob = travel_probs[[c]]),
                   character(1))

  # 9. Child characteristics
  sex    <- rbinom(N, 1, 0.4769)
  bw_cat <- sample(c("<2.5", "2.5-4", ">4"), N, replace = TRUE,
                   prob = c(0.034, 0.911, 0.055))
  birth_weight <- ifelse(bw_cat == "<2.5",  rnorm(N, 2.3, 0.20),
                  ifelse(bw_cat == "2.5-4", rnorm(N, 3.2, 0.35),
                                             rnorm(N, 4.3, 0.25)))

  # 10. Baseline HAZ (country-specific)
  haz_mean <- c(Uganda = -0.10, Malawi = -0.20,
                Zimbabwe =  0.15, "South Africa" =  0.00)
  haz_sd   <- c(Uganda =  1.25, Malawi =  1.30,
                Zimbabwe =  1.15, "South Africa" =  1.20)
  haz0 <- rnorm(N, haz_mean[country], haz_sd[country])

  # 11. Random effects
  Sigma_b <- matrix(c(sigma_b0^2,
                      rho_b01 * sigma_b0 * sigma_b1,
                      rho_b01 * sigma_b0 * sigma_b1,
                      sigma_b1^2), nrow = 2)
  re  <- mvrnorm(N, mu = c(0, 0), Sigma = Sigma_b)
  b0  <- re[, 1]
  b1  <- re[, 2]

  # 12. Longitudinal structure — vectorised
  time_pts <- c(0, 6, 12, 18, 24, 36, 48, 60)
  T        <- length(time_pts)

  idx   <- rep(seq_len(N), each  = T)
  t_vec <- rep(time_pts,   times = N)

  g1 <- t_vec
  g2 <- log(t_vec + 1)
  g3 <- 1 / (t_vec + 1)

  # True HAZ fixed effects (Malawi reference)
  country_effect <- ifelse(country[idx] == "South Africa",  0.45,
                    ifelse(country[idx] == "Uganda",          0.20,
                    ifelse(country[idx] == "Zimbabwe",        0.15, 0.00)))

  mu <- (-0.50)  +
        (-0.015) * g1 +
          0.30   * g2 +
        (-0.10)  * g3 +
        country_effect +
        b0[idx] +
        b1[idx] * g1

  # CHANGED: uses the sigma_eps argument instead of the hard-coded 0.80
  Y <- mu + rnorm(N * T, 0, sigma_eps)
  Y[abs(Y) > 4] <- NA   # HAZ outside ±4 treated as implausible

  df <- data.frame(
    child_id     = idx,
    country      = country[idx],
    maternal_age = maternal_age[idx],
    education    = edu_cat[idx],
    edu_high     = edu_high[idx],
    work         = work[idx],
    marital      = marital[idx],
    electricity  = electricity[idx],
    tapwater     = tapwater[idx],
    toilet       = toilet[idx],
    travel       = travel[idx],
    sex          = sex[idx],
    birth_weight = birth_weight[idx],
    age_mo       = t_vec,
    g1           = g1,
    g2           = g2,
    g3           = g3,
    outcome      = Y,
    stringsAsFactors = FALSE
  )

  # 13. MAR dropout
  ids       <- unique(df$child_id)
  drop_ids  <- sample(ids, size = round(dropout * N), replace = FALSE)
  drop_time <- sample(2:T, size = length(drop_ids), replace = TRUE)
  for (k in seq_along(drop_ids)) {
    id   <- drop_ids[k]
    tmin <- time_pts[drop_time[k]]
    df$outcome[df$child_id == id & df$age_mo >= tmin] <- NA
  }

  return(df)
}
```

---

# 2. Model-Fitting Function

```{r chunk-2-fit-model}
fit_LMM_HAZ <- function(dat) {
  dat$country <- relevel(factor(dat$country), ref = "Malawi")
  ctrl        <- lmerControl(optimizer = "bobyqa",
                              optCtrl  = list(maxfun = 2e5))
  formula_lmm <- outcome ~ g1 + g2 + g3 + country + (1 + g1 | child_id)
  fit <- tryCatch(
    lmer(formula_lmm, data = dat, REML = FALSE, control = ctrl),
    warning = function(w) {
      m <- suppressWarnings(
        lmer(formula_lmm, data = dat, REML = FALSE, control = ctrl)
      )
      attr(m, "converge_warning") <- conditionMessage(w)
      m
    },
    error = function(e) NULL
  )
  return(fit)
}
```

---

# 3. Results-Extraction Function — EXPANDED

```{r chunk-3-extract}
# CHANGED: now also extracts std.error so downstream functions can compute
# empirical coverage and Type I error rate.
extract_results <- function(model) {
  if (is.null(model)) return(NULL)
  converged <- is.null(attr(model, "converge_warning")) && !isSingular(model)
  broom.mixed::tidy(model, effects = "fixed") %>%
    dplyr::select(term, estimate, std.error, statistic, p.value) %>%
    mutate(converged = converged)
}
```

---

# 4. Single-Replication Wrapper

```{r chunk-4-single-sim}
# CHANGED: sigma_eps argument forwarded to generate_HAZ_data().
single_sim_HAZ <- function(N         = 500,
                            sigma_b0  = 0.6,
                            sigma_b1  = 0.02,
                            rho_b01   = 0.2,
                            sigma_eps = 0.80) {
  dat   <- generate_HAZ_data(N         = N,
                              sigma_b0  = sigma_b0,
                              sigma_b1  = sigma_b1,
                              rho_b01   = rho_b01,
                              sigma_eps = sigma_eps)
  model <- fit_LMM_HAZ(dat)
  extract_results(model)
}
```

---

# 5. Simulation Runner

```{r chunk-5-run-sim}
# CHANGED: sigma_eps argument forwarded.
run_simulation_HAZ <- function(R         = 200,
                                N         = 500,
                                sigma_b0  = 0.6,
                                sigma_b1  = 0.02,
                                rho_b01   = 0.2,
                                sigma_eps = 0.80) {
  res_list <- pblapply(seq_len(R), function(i)
    single_sim_HAZ(N         = N,
                   sigma_b0  = sigma_b0,
                   sigma_b1  = sigma_b1,
                   rho_b01   = rho_b01,
                   sigma_eps = sigma_eps)
  )
  rbindlist(res_list, idcol = "rep", fill = TRUE)
}
```

---

# 6. True HAZ Parameter Vector

```{r chunk-6-true-values}
true_values <- c(
  "(Intercept)"         = -0.50,
  "g1"                  = -0.015,
  "g2"                  =  0.30,
  "g3"                  = -0.10,
  "countrySouth Africa" =  0.45,
  "countryUganda"       =  0.20,
  "countryZimbabwe"     =  0.15
)
```

---

# 7. Performance-Summary Helper — EXPANDED

```{r chunk-7-performance}
# CHANGED: now computes four metrics for each term:
#   1. Bias             — systematic deviation
#   2. RMSE             — combined bias + variance
#   3. Coverage (95%)   — % of 95% CIs that contain β_true
#   4. Convergence rate — % of reps where the LMM converged
#
# NOTE on Type I error:
#   Type I error is meaningful only when the true effect equals zero.
#   None of the parameters in `true_values` are zero in the main scenarios.
#   For the null-scenario check (Section 8c), country effects ARE set to zero
#   and Type I error is computed directly in that section rather than inside
#   this helper — keeping the helper general-purpose.
summarise_performance <- function(sim_results, label = "") {

  # Convergence rate: proportion of ALL reps (not just converged)
  conv_rate <- sim_results %>%
    group_by(term) %>%
    summarise(convergence_rate = mean(converged, na.rm = TRUE), .groups = "drop")

  # Main metrics restricted to converged reps
  main <- sim_results %>%
    filter(converged == TRUE) %>%
    group_by(term) %>%
    summarise(
      n_reps        = n(),
      mean_estimate = mean(estimate,   na.rm = TRUE),
      bias          = mean(estimate  - true_values[term], na.rm = TRUE),
      rmse          = sqrt(mean((estimate - true_values[term])^2, na.rm = TRUE)),
      sd_estimate   = sd(estimate,     na.rm = TRUE),
      # Coverage: 95% CI = estimate ± 1.96 * SE
      coverage_95   = mean(
        abs(estimate - true_values[term]) <= 1.96 * std.error,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%
    left_join(conv_rate, by = "term") %>%
    mutate(scenario = label)

  return(main)
}
```

---

# 8. Baseline Simulation *(N = 736)*

## 8a. Run Baseline

```{r chunk-8a-baseline-sim, cache=TRUE}
# NOTE: R_REPS_MAIN = 1000 — first run will take ~5–10 minutes depending on hardware.
# All subsequent runs load instantly from cache.
sim_base_HAZ <- {
  cached <- cache_load("sim_base_HAZ_k1000")
  if (!is.null(cached)) {
    message("Loaded HAZ baseline simulation from cache.")
    cached
  } else {
    message("Running HAZ baseline simulation (N = ", N_TARGET,
            ", R = ", R_REPS_MAIN, " reps) ...")
    res <- run_simulation_HAZ(R = R_REPS_MAIN, N = N_TARGET)
    cache_save(res, "sim_base_HAZ_k1000")
  }
}

perf_base_HAZ <- summarise_performance(
  sim_base_HAZ,
  label = paste0("HAZ Baseline — Real Study Anchor (N = ", N_TARGET, ")")
)

knitr::kable(
  perf_base_HAZ %>%
    dplyr::select(term, n_reps, mean_estimate, bias, rmse,
                  coverage_95, convergence_rate, scenario),
  digits  = 4,
  caption = paste0(
    "HAZ baseline performance (N = ", N_TARGET,
    " — real study size; reference country = Malawi; R = ", R_REPS_MAIN, " reps)"
  )
)
```

## 8b. Coverage Plot *(baseline)*

```{r chunk-8b-coverage-plot, fig.width=9, fig.height=5}
# Nominal coverage = 0.95; the shaded band shows ±2 × Monte Carlo SE.
# With R = 1000, MC-SE = sqrt(0.95 * 0.05 / 1000) ≈ 0.007.
mc_se <- sqrt(0.95 * 0.05 / R_REPS_MAIN)

perf_base_HAZ %>%
  filter(term %in% names(true_values)) %>%
  ggplot(aes(x = term, y = coverage_95)) +
  geom_point(size = 3, colour = "steelblue") +
  geom_errorbar(aes(ymin = coverage_95 - 2 * mc_se,
                    ymax = coverage_95 + 2 * mc_se),
                width = 0.2, colour = "steelblue") +
  geom_hline(yintercept = 0.95,  linetype = "dashed", colour = "red") +
  geom_hline(yintercept = 0.95 - 2 * mc_se, linetype = "dotted", colour = "grey50") +
  geom_hline(yintercept = 0.95 + 2 * mc_se, linetype = "dotted", colour = "grey50") +
  annotate("rect",
           xmin = -Inf, xmax = Inf,
           ymin = 0.95 - 2 * mc_se,
           ymax = 0.95 + 2 * mc_se,
           alpha = 0.10, fill = "red") +
  scale_y_continuous(limits = c(0.85, 1.00),
                     labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "HAZ: Empirical Coverage of 95% Wald CIs — Baseline (N = 736)",
    subtitle = paste0("Red dashed = nominal 95%; shaded band = ±2 × MC-SE (",
                      R_REPS_MAIN, " reps); error bars = ±2 × MC-SE per term"),
    x        = "Fixed-effect term",
    y        = "Empirical coverage"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## 8c. Type I Error — Null Country-Effect Scenario

```{r chunk-8c-typeI, cache=TRUE}
# Type I error is defined only when the true effect = 0.
# Here we generate data with ALL country effects set to 0 (null scenario)
# and check how often the LMM incorrectly rejects H₀: β_country = 0.
#
# The data-generating model is otherwise identical to baseline (N = 736).
# We override country_effect inside a local wrapper to ensure β_country = 0.

generate_HAZ_null <- function(N         = 736,
                               dropout   = 0.20,
                               sigma_b0  = 0.6,
                               sigma_b1  = 0.02,
                               rho_b01   = 0.2,
                               sigma_eps = 0.80) {
  df <- generate_HAZ_data(N         = N,
                           dropout   = dropout,
                           sigma_b0  = sigma_b0,
                           sigma_b1  = sigma_b1,
                           rho_b01   = rho_b01,
                           sigma_eps = sigma_eps)
  # Recompute outcome with zero country effects (all else unchanged)
  g1 <- df$g1; g2 <- df$g2; g3 <- df$g3
  mu_null <- (-0.50) + (-0.015)*g1 + 0.30*g2 + (-0.10)*g3
  # Re-draw residuals at σ_ε = sigma_eps; random effects already baked in via Y
  # Simpler: just zero out the country contribution from the stored outcome
  country_effect_orig <- ifelse(df$country == "South Africa",  0.45,
                          ifelse(df$country == "Uganda",         0.20,
                          ifelse(df$country == "Zimbabwe",       0.15, 0.00)))
  df$outcome <- df$outcome - country_effect_orig
  return(df)
}

run_typeI_HAZ <- function(R = R_REPS_MAIN) {
  pblapply(seq_len(R), function(i) {
    dat   <- generate_HAZ_null(N = N_TARGET)
    model <- fit_LMM_HAZ(dat)
    if (is.null(model)) return(NULL)
    broom.mixed::tidy(model, effects = "fixed") %>%
      dplyr::select(term, p.value) %>%
      mutate(converged = is.null(attr(model, "converge_warning")) &&
                           !isSingular(model))
  }) %>%
    rbindlist(idcol = "rep", fill = TRUE)
}

typeI_results <- {
  cached <- cache_load("HAZ_typeI_k1000")
  if (!is.null(cached)) { message("Loaded Type I results from cache."); cached }
  else {
    message("Running Type I error simulation (R = ", R_REPS_MAIN, ") ...")
    res <- run_typeI_HAZ(R = R_REPS_MAIN)
    cache_save(res, "HAZ_typeI_k1000")
  }
}

typeI_table <- typeI_results %>%
  filter(converged, term %in% c("countrySouth Africa",
                                 "countryUganda", "countryZimbabwe")) %>%
  group_by(term) %>%
  summarise(
    n_reps      = n(),
    typeI_error = mean(p.value < 0.05, na.rm = TRUE),
    .groups     = "drop"
  )

knitr::kable(
  typeI_table,
  digits  = 4,
  caption = paste0(
    "Type I error rate for country fixed effects under the null (β_country = 0); ",
    "nominal α = 0.05; R = ", R_REPS_MAIN
  )
)
```

---

# 9. Sample-Size Scenarios

```{r chunk-9-n-scenarios, cache=TRUE}
sample_size_scenarios <- list(
  list(N = round(N_TARGET * 0.5),
       label = "Half the Real Study (N = 368)"),
  list(N = N_TARGET,
       label = "Real Study — Anchor (N = 736)"),
  list(N = N_TARGET * 2,
       label = "Double the Real Study (N = 1 472)"),
  list(N = N_TARGET * 5,
       label = "Registry Scale — Five Times Real (N = 3 680)")
)

n_scenario_results_HAZ <- lapply(sample_size_scenarios, function(sc) {
  cname  <- paste0("HAZ_n_scenario_N", sc$N, "_k1000")
  cached <- cache_load(cname)
  if (!is.null(cached)) {
    message("Loaded HAZ '", sc$label, "' from cache.")
    return(mutate(cached, N = sc$N, n_label = sc$label))
  }
  cat("\n▶ HAZ — Running sample-size scenario:", sc$label, "\n")
  res <- run_simulation_HAZ(R = R_REPS_MAIN, N = sc$N) %>%
    mutate(N = sc$N, n_label = sc$label)
  cache_save(res, cname)
  res
})

n_scenario_df_HAZ <- rbindlist(n_scenario_results_HAZ, fill = TRUE) %>%
  mutate(n_label = factor(
    n_label,
    levels = sapply(sample_size_scenarios, `[[`, "label")
  ))

perf_n_HAZ <- n_scenario_df_HAZ %>%
  group_by(N, n_label) %>%
  group_modify(~summarise_performance(.x, label = as.character(.y$n_label))) %>%
  ungroup()

knitr::kable(
  perf_n_HAZ %>%
    dplyr::select(N, n_label, term, n_reps, bias, rmse,
                  coverage_95, convergence_rate),
  digits  = 4,
  caption = paste0(
    "HAZ performance across sample-size scenarios (anchor = N = ", N_TARGET,
    "; reference country = Malawi; R = ", R_REPS_MAIN, " reps)"
  )
)
```

---

# 10. Random-Effects Scenarios

## 10a. Define the 16-Scenario Grid *(Groups A–D)*

```{r chunk-10a-re-define}
# ── Groups A, B, C are unchanged from the original script ────────────────────
# ── Group D (NEW): Worst-Case Combined scenarios ──────────────────────────────
#
#   Rationale for "worst-case" design:
#   The hardest conditions for an LMM arise when:
#     (a) Between-child variance is extreme (σ₀ and σ₁ both large), AND
#     (b) Intercept–slope correlation is strong (|ρ| near 1),
#   because the random-effects covariance matrix becomes near-singular.
#   These combined scenarios test whether the LMM remains well-calibrated
#   (unbiased, nominal coverage) under realistic worst-case assumptions
#   drawn from the observed HAZ literature.
#
#   Three sub-scenarios:
#     D1. Extreme variance + strong negative correlation
#         (children with high HAZ at birth lose height-gain advantage)
#     D2. Extreme variance + strong positive correlation
#         (children with high HAZ at birth maintain/widen advantage)
#     D3. Near-maximal stress: σ₀ = 2.0, σ₁ = 0.12, ρ = −0.90
#         (pathological but informative boundary case)
# ─────────────────────────────────────────────────────────────────────────────
re_scenarios <- list(

  # ── Group A: Variance Magnitude ───────────────────────────────────────────
  "Minimal Between-Child Variation" = list(
    sigma_b0 = 0.20, sigma_b1 = 0.005, rho_b01 =  0.0,
    group    = "A: Variance Magnitude",
    label    = "Minimal Between-Child Variation\n(σ₀=0.20, σ₁=0.005)"
  ),
  "Low Between-Child Variation" = list(
    sigma_b0 = 0.40, sigma_b1 = 0.010, rho_b01 =  0.0,
    group    = "A: Variance Magnitude",
    label    = "Low Between-Child Variation\n(σ₀=0.40, σ₁=0.010)"
  ),
  "Baseline Variation — Study Anchor \u2605" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 =  0.2,
    group    = "A: Variance Magnitude",
    label    = "Baseline Variation — Study Anchor \u2605\n(σ₀=0.60, σ₁=0.020)"
  ),
  "High Between-Child Variation" = list(
    sigma_b0 = 1.00, sigma_b1 = 0.050, rho_b01 =  0.2,
    group    = "A: Variance Magnitude",
    label    = "High Between-Child Variation\n(σ₀=1.00, σ₁=0.050)"
  ),
  "Extreme Between-Child Variation" = list(
    sigma_b0 = 1.50, sigma_b1 = 0.080, rho_b01 =  0.2,
    group    = "A: Variance Magnitude",
    label    = "Extreme Between-Child Variation\n(σ₀=1.50, σ₁=0.080)"
  ),

  # ── Group B: Slope Variability ────────────────────────────────────────────
  "Near-Identical Growth Trajectories" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.005, rho_b01 =  0.2,
    group    = "B: Slope Variability",
    label    = "Near-Identical Growth Trajectories\n(σ₁=0.005)"
  ),
  "Moderately Diverging Growth Trajectories" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.040, rho_b01 =  0.2,
    group    = "B: Slope Variability",
    label    = "Moderately Diverging Growth Trajectories\n(σ₁=0.040)"
  ),
  "Highly Diverging Growth Trajectories" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.100, rho_b01 =  0.2,
    group    = "B: Slope Variability",
    label    = "Highly Diverging Growth Trajectories\n(σ₁=0.100)"
  ),

  # ── Group C: Intercept–Slope Correlation ──────────────────────────────────
  "Strong Negative Intercept-Slope Correlation" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 = -0.7,
    group    = "C: Intercept-Slope Correlation",
    label    = "Strong Negative Intercept-Slope Correlation\n(rho=-0.70)"
  ),
  "Moderate Negative Intercept-Slope Correlation" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 = -0.4,
    group    = "C: Intercept-Slope Correlation",
    label    = "Moderate Negative Intercept-Slope Correlation\n(rho=-0.40)"
  ),
  "No Intercept-Slope Correlation" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 =  0.0,
    group    = "C: Intercept-Slope Correlation",
    label    = "No Intercept-Slope Correlation\n(rho=0.00)"
  ),
  "Moderate Positive Intercept-Slope Correlation" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 =  0.4,
    group    = "C: Intercept-Slope Correlation",
    label    = "Moderate Positive Intercept-Slope Correlation\n(rho=+0.40)"
  ),
  "Strong Positive Intercept-Slope Correlation" = list(
    sigma_b0 = 0.60, sigma_b1 = 0.020, rho_b01 =  0.7,
    group    = "C: Intercept-Slope Correlation",
    label    = "Strong Positive Intercept-Slope Correlation\n(rho=+0.70)"
  ),

  # ── Group D: Worst-Case Combined (NEW) ────────────────────────────────────
  # Combines extreme variance (Group A ceiling) with strong correlation
  # (Group C extremes) to create the most challenging scenario for the LMM.
  # Expected outcomes: reduced convergence, inflated SE, possible coverage drop.
  "Worst-Case: Extreme Variance + Strong Negative Correlation" = list(
    sigma_b0 = 1.50, sigma_b1 = 0.080, rho_b01 = -0.70,
    group    = "D: Worst-Case Combined",
    label    = "Worst-Case: Extreme Var + Strong Neg Corr\n(σ₀=1.50, σ₁=0.080, rho=-0.70)"
  ),
  "Worst-Case: Extreme Variance + Strong Positive Correlation" = list(
    sigma_b0 = 1.50, sigma_b1 = 0.080, rho_b01 =  0.70,
    group    = "D: Worst-Case Combined",
    label    = "Worst-Case: Extreme Var + Strong Pos Corr\n(σ₀=1.50, σ₁=0.080, rho=+0.70)"
  ),
  "Worst-Case: Near-Maximal Stress" = list(
    sigma_b0 = 2.00, sigma_b1 = 0.120, rho_b01 = -0.90,
    group    = "D: Worst-Case Combined",
    label    = "Worst-Case: Near-Maximal Stress\n(σ₀=2.00, σ₁=0.120, rho=-0.90)"
  )
)

# Parameter look-up table
re_param_table <- imap_dfr(re_scenarios, function(sc, nm) {
  data.frame(
    scenario_name = nm,
    group         = sc$group,
    sigma_b0      = sc$sigma_b0,
    sigma_b1      = sc$sigma_b1,
    rho_b01       = sc$rho_b01,
    var_b0        = sc$sigma_b0^2,
    var_b1        = sc$sigma_b1^2,
    cov_b01       = sc$rho_b01 * sc$sigma_b0 * sc$sigma_b1
  )
})

knitr::kable(re_param_table, digits = 4,
             caption = "HAZ random-effects scenario parameter grid (star = Study Anchor)")
```

## 10b. Run RE Scenarios

```{r chunk-10b-re-scenarios-run, cache=TRUE}
re_results_HAZ <- imap(re_scenarios, function(sc, name) {
  safe_name <- gsub("[^A-Za-z0-9]", "_", name)
  cname     <- paste0("HAZ_re_scenario_", safe_name, "_k1000")
  cached    <- cache_load(cname)
  if (!is.null(cached)) {
    message("Loaded HAZ RE scenario '", name, "' from cache.")
    return(mutate(cached, re_scenario = sc$label, re_group = sc$group))
  }
  cat("\n▶ HAZ — Running RE scenario:", name, "\n")
  res <- run_simulation_HAZ(
    R        = R_REPS_MAIN,
    N        = N_TARGET,
    sigma_b0 = sc$sigma_b0,
    sigma_b1 = sc$sigma_b1,
    rho_b01  = sc$rho_b01
  ) %>%
    mutate(re_scenario = sc$label, re_group = sc$group)
  cache_save(res, cname)
  res
})

re_results_df_HAZ <- rbindlist(re_results_HAZ, fill = TRUE)
```

## 10c. RE Performance Table

```{r chunk-10c-re-performance}
key_terms <- c(
  "(Intercept)", "g1", "g2", "g3",
  "countrySouth Africa", "countryUganda", "countryZimbabwe"
)

perf_re_HAZ <- re_results_df_HAZ %>%
  group_by(re_group, re_scenario) %>%
  group_modify(~summarise_performance(.x, label = .y$re_scenario)) %>%
  ungroup()

knitr::kable(
  perf_re_HAZ %>%
    dplyr::select(re_group, re_scenario, term, n_reps,
                  bias, rmse, coverage_95, convergence_rate),
  digits  = 4,
  caption = paste0(
    "HAZ fixed-effect performance across RE scenarios — Groups A–D ",
    "(N = ", N_TARGET, "; R = ", R_REPS_MAIN, " reps; reference country = Malawi)"
  )
)
```

---

# 11. Residual Variance Scenarios *(Group E — NEW)*

```{r chunk-11-resid-var, cache=TRUE}
# ── Design rationale ──────────────────────────────────────────────────────────
# σ_ε captures within-child measurement noise (anthropometry error, day-to-day
# biological variation, data entry error).  The default σ_ε = 0.80 reflects
# typical HAZ reproducibility.  Scenarios range from "near-perfect measurement"
# (σ_ε = 0.40 — akin to DXA / digital stadiometers with repeat measures) to
# "high noise" (σ_ε = 1.60 — field conditions with single-observer HAZ).
#
# A key question: does inflated residual variance primarily inflate SE (and
# thus widen CIs, maintaining coverage) or does it also bias fixed effects?
# Under a correctly specified LMM the answer should be "SE only" — deviations
# from nominal coverage signal mis-specification or small-sample issues.
# ─────────────────────────────────────────────────────────────────────────────
resid_scenarios <- list(
  list(sigma_eps = 0.40, label = "Low Residual Noise (σ_ε = 0.40)"),
  list(sigma_eps = 0.60, label = "Moderate-Low Residual Noise (σ_ε = 0.60)"),
  list(sigma_eps = 0.80, label = "Baseline Residual Noise ★ (σ_ε = 0.80)"),
  list(sigma_eps = 1.00, label = "Moderate-High Residual Noise (σ_ε = 1.00)"),
  list(sigma_eps = 1.20, label = "High Residual Noise (σ_ε = 1.20)"),
  list(sigma_eps = 1.60, label = "Very High Residual Noise (σ_ε = 1.60)")
)

resid_results_HAZ <- lapply(resid_scenarios, function(sc) {
  safe  <- gsub("[^A-Za-z0-9]", "_", sc$label)
  cname <- paste0("HAZ_resid_", safe, "_k1000")
  cached <- cache_load(cname)
  if (!is.null(cached)) {
    message("Loaded residual-variance scenario '", sc$label, "' from cache.")
    return(mutate(cached, eps_label = sc$label, sigma_eps = sc$sigma_eps))
  }
  cat("\n▶ HAZ — Residual variance scenario:", sc$label, "\n")
  res <- run_simulation_HAZ(
    R         = R_REPS_MAIN,
    N         = N_TARGET,
    sigma_eps = sc$sigma_eps
  ) %>%
    mutate(eps_label = sc$label, sigma_eps = sc$sigma_eps)
  cache_save(res, cname)
  res
})

resid_df_HAZ <- rbindlist(resid_results_HAZ, fill = TRUE) %>%
  mutate(eps_label = factor(eps_label,
                             levels = sapply(resid_scenarios, `[[`, "label")))

perf_resid_HAZ <- resid_df_HAZ %>%
  group_by(sigma_eps, eps_label) %>%
  group_modify(~summarise_performance(.x, label = as.character(.y$eps_label))) %>%
  ungroup()

knitr::kable(
  perf_resid_HAZ %>%
    dplyr::select(sigma_eps, eps_label, term, n_reps,
                  bias, rmse, coverage_95, convergence_rate),
  digits  = 4,
  caption = paste0(
    "HAZ performance across residual-variance scenarios (Group E) — ",
    "N = ", N_TARGET, "; R = ", R_REPS_MAIN, " reps; reference = Malawi"
  )
)
```

---

# 12. Crossed Scenarios *(N × RE)*

```{r chunk-12-crossed, cache=TRUE}
crossed_n_values <- c(
  round(N_TARGET * 0.5),
  N_TARGET,
  N_TARGET * 2
)

crossed_n_labels <- c(
  paste0("Half the Real Study (N = ",    round(N_TARGET * 0.5), ")"),
  paste0("Real Study — Anchor (N = ",    N_TARGET, ")"),
  paste0("Double the Real Study (N = ",  N_TARGET * 2, ")")
)
names(crossed_n_labels) <- as.character(crossed_n_values)

crossed_grid <- expand.grid(
  N       = crossed_n_values,
  re_name = names(re_scenarios),
  stringsAsFactors = FALSE
)

crossed_results_HAZ <- pmap_dfr(crossed_grid, function(N, re_name) {
  sc        <- re_scenarios[[re_name]]
  safe_name <- gsub("[^A-Za-z0-9]", "_", re_name)
  cname     <- paste0("HAZ_crossed_N", N, "_", safe_name, "_k200")
  cached    <- cache_load(cname)
  if (!is.null(cached)) {
    message("Loaded HAZ crossed cell: '",
            crossed_n_labels[as.character(N)], "' x '", re_name, "'")
    return(mutate(cached,
                  N           = N,
                  n_label     = crossed_n_labels[as.character(N)],
                  re_scenario = sc$label,
                  re_group    = sc$group))
  }
  cat("\n▶ HAZ crossed:", crossed_n_labels[as.character(N)], "x", re_name, "\n")
  res <- run_simulation_HAZ(
    R        = R_REPS_CROSS,
    N        = N,
    sigma_b0 = sc$sigma_b0,
    sigma_b1 = sc$sigma_b1,
    rho_b01  = sc$rho_b01
  ) %>%
    mutate(N           = N,
           n_label     = crossed_n_labels[as.character(N)],
           re_scenario = sc$label,
           re_group    = sc$group)
  cache_save(res, cname)
  res
})

perf_crossed_HAZ <- crossed_results_HAZ %>%
  group_by(N, n_label, re_group, re_scenario) %>%
  group_modify(~summarise_performance(
    .x,
    label = paste0(.y$n_label, " | ", .y$re_scenario)
  )) %>%
  ungroup()

knitr::kable(
  perf_crossed_HAZ %>%
    dplyr::select(N, n_label, re_group, re_scenario, term,
                  bias, rmse, coverage_95, convergence_rate),
  digits  = 4,
  caption = "HAZ crossed (N × RE) performance table (reference country = Malawi)"
)
```

---

# 13. Visualisation

## 13a. Bias by RE Scenario *(Groups A–D, faceted)*

```{r chunk-13a-bias-plot, fig.width = 13, fig.height = 9}
perf_re_HAZ %>%
  filter(term %in% key_terms) %>%
  mutate(re_scenario = gsub("\n", " ", re_scenario)) %>%
  ggplot(aes(x = re_scenario, y = bias, fill = re_group)) +
  geom_col(show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(term ~ re_group, scales = "free") +
  labs(
    title    = "HAZ: Bias of Fixed-Effect Estimates Across RE Scenarios (Groups A–D)",
    subtitle = paste0("N = ", N_TARGET, "; R = ", R_REPS_MAIN,
                      " reps; dashed = zero bias; Group D = worst-case combined"),
    x        = "Random-Effects Scenario",
    y        = "Bias (estimated − true)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7))
```

## 13b. Coverage by RE Scenario

```{r chunk-13b-coverage-re, fig.width = 13, fig.height = 7}
mc_se <- sqrt(0.95 * 0.05 / R_REPS_MAIN)

perf_re_HAZ %>%
  filter(term %in% key_terms) %>%
  mutate(re_scenario = gsub("\n", " ", re_scenario),
         worst_case  = grepl("Worst-Case", re_scenario)) %>%
  ggplot(aes(x = re_scenario, y = coverage_95,
             colour = re_group, shape = worst_case)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = coverage_95 - 2 * mc_se,
                    ymax = coverage_95 + 2 * mc_se),
                width = 0.3) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
  annotate("rect",
           xmin = -Inf, xmax = Inf,
           ymin = 0.95 - 2 * mc_se,
           ymax = 0.95 + 2 * mc_se,
           alpha = 0.08, fill = "red") +
  facet_wrap(~term, scales = "free_x", ncol = 2) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0.80, 1.00)) +
  scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 17),
                     labels = c("Standard", "Worst-Case (Group D)")) +
  labs(
    title    = "HAZ: Empirical Coverage of 95% CIs Across RE Scenarios",
    subtitle = paste0("Triangles = worst-case (Group D); red band = ±2 × MC-SE (",
                      R_REPS_MAIN, " reps)"),
    x        = "RE Scenario",
    y        = "Empirical Coverage",
    colour   = "RE Group",
    shape    = "Scenario type"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 6),
        legend.position = "bottom")
```

## 13c. RMSE by Sample Size

```{r chunk-13c-rmse-n, fig.width = 11, fig.height = 6}
n_label_order <- sapply(sample_size_scenarios, `[[`, "label")

perf_n_HAZ %>%
  filter(term %in% key_terms) %>%
  mutate(n_label = factor(n_label, levels = n_label_order)) %>%
  ggplot(aes(x = n_label, y = rmse, group = term, colour = term)) +
  geom_line() +
  geom_point(size = 2) +
  geom_vline(
    xintercept = which(n_label_order == paste0("Real Study — Anchor (N = ", N_TARGET, ")")),
    linetype = "dashed", colour = "grey40"
  ) +
  labs(
    title    = "HAZ: RMSE by Sample-Size Scenario",
    subtitle = paste0("Dashed = real study anchor (N = ", N_TARGET,
                      "); R = ", R_REPS_MAIN, " reps"),
    x        = "Sample-Size Scenario",
    y        = "RMSE",
    colour   = "Fixed-Effect Term"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1, size = 9))
```

## 13d. Coverage by Sample Size

```{r chunk-13d-coverage-n, fig.width = 11, fig.height = 6}
mc_se <- sqrt(0.95 * 0.05 / R_REPS_MAIN)

perf_n_HAZ %>%
  filter(term %in% key_terms) %>%
  mutate(n_label = factor(n_label, levels = n_label_order)) %>%
  ggplot(aes(x = n_label, y = coverage_95, group = term, colour = term)) +
  geom_line() +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
  annotate("rect",
           xmin = -Inf, xmax = Inf,
           ymin = 0.95 - 2 * mc_se,
           ymax = 0.95 + 2 * mc_se,
           alpha = 0.08, fill = "red") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0.85, 1.00)) +
  labs(
    title    = "HAZ: 95% CI Coverage by Sample-Size Scenario",
    subtitle = "Red dashed = nominal 95%; shaded band = ±2 × MC-SE",
    x        = "Sample-Size Scenario",
    y        = "Empirical Coverage",
    colour   = "Fixed-Effect Term"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1, size = 9))
```

## 13e. Residual Variance — RMSE & Coverage

```{r chunk-13e-resid-plots, fig.width = 12, fig.height = 10}
png("HAZ Residual Variance.png", width = 800, height = 600)
p_rmse_eps <- perf_resid_HAZ %>%
  filter(term %in% key_terms) %>%
  ggplot(aes(x = sigma_eps, y = rmse, colour = term)) +
  geom_line() + geom_point(size = 2) +
  geom_vline(xintercept = 0.80, linetype = "dashed", colour = "grey40") +
  annotate("text", x = 0.80, y = Inf,
           label = "Baseline\n(σ_ε=0.80)", vjust = 1.5, hjust = -0.1,
           size = 3, colour = "grey40") +
  labs(title = "RMSE vs Residual Noise Level",
       x = "Residual SD (σ_ε)", y = "RMSE", colour = "Term") +
  theme_bw()

mc_se <- sqrt(0.95 * 0.05 / R_REPS_MAIN)

p_cov_eps <- perf_resid_HAZ %>%
  filter(term %in% key_terms) %>%
  ggplot(aes(x = sigma_eps, y = coverage_95, colour = term)) +
  geom_line() + geom_point(size = 2) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
  annotate("rect",
           xmin = -Inf, xmax = Inf,
           ymin = 0.95 - 2 * mc_se,
           ymax = 0.95 + 2 * mc_se,
           alpha = 0.08, fill = "red") +
  geom_vline(xintercept = 0.80, linetype = "dashed", colour = "grey40") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0.85, 1.00)) +
  labs(title = "95% CI Coverage vs Residual Noise Level",
       x = "Residual SD (σ_ε)", y = "Empirical Coverage", colour = "Term") +
  theme_bw()

gridExtra::grid.arrange(p_rmse_eps, p_cov_eps, nrow = 2,
  top = grid::textGrob(
    paste0("HAZ: Impact of Residual Variance on RMSE and Coverage\n",
           "(Group E scenarios; N = ", N_TARGET,
           "; R = ", R_REPS_MAIN, " reps; reference = Malawi)"),
    gp = grid::gpar(fontsize = 13, fontface = "bold")
  )
)
dev.off()
```

## 13f. Empirical Distribution of Estimates *(baseline)*

```{r chunk-13f-density-plot, fig.width = 11, fig.height = 7}
truth_df <- data.frame(
  term  = names(true_values),
  truth = true_values
) %>% filter(term %in% key_terms)

sim_base_HAZ %>%
  filter(converged == TRUE, term %in% key_terms) %>%
  ggplot(aes(x = estimate)) +
  geom_density(fill = "darkorange", alpha = 0.45) +
  geom_vline(data = truth_df, aes(xintercept = truth),
             colour = "red", linetype = "dashed") +
  facet_wrap(~term, scales = "free") +
  labs(
    title    = "HAZ: Empirical Sampling Distribution of Fixed-Effect Estimates",
    subtitle = paste0("Baseline: N = ", N_TARGET, ", R = ", R_REPS_MAIN,
                      " reps; red dashed = true value"),
    x = "Estimate", y = "Density"
  ) +
  theme_bw()
```

## 13g. Bias Heat-Map *(crossed N × RE grid)*

```{r chunk-13g-bias-heatmap, fig.width = 12, fig.height = 16}
heatmap_term <- "g2"

perf_crossed_HAZ %>%
  filter(term == heatmap_term) %>%
  mutate(
    re_scenario = gsub("\n", " ", re_scenario),
    n_label     = factor(n_label,
                         levels = sapply(sample_size_scenarios[1:3], `[[`, "label"))
  ) %>%
  ggplot(aes(x = n_label, y = re_scenario, fill = bias)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = round(bias, 3)), size = 3) +
  scale_fill_gradient2(low = "steelblue", mid = "white",
                       high = "firebrick", midpoint = 0) +
  facet_wrap(~re_group, scales = "free_y", ncol = 1) +
  labs(
    title    = "HAZ Bias Heat-Map: Sample Size × RE Scenarios (Groups A–D)",
    subtitle = paste0("Term: ", heatmap_term,
                      "; Group D = worst-case combined; reference = Malawi"),
    x = "Sample-Size Scenario", y = "RE Scenario", fill = "Bias"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8))
```

## 13h. Coverage Heat-Map *(crossed N × RE grid — NEW)*

```{r chunk-13h-coverage-heatmap, fig.width = 12, fig.height = 16}
perf_crossed_HAZ %>%
  filter(term == heatmap_term) %>%
  mutate(
    re_scenario = gsub("\n", " ", re_scenario),
    n_label     = factor(n_label,
                         levels = sapply(sample_size_scenarios[1:3], `[[`, "label"))
  ) %>%
  ggplot(aes(x = n_label, y = re_scenario, fill = coverage_95)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = scales::percent(coverage_95, accuracy = 1)), size = 3) +
  scale_fill_gradient2(low = "firebrick", mid = "white",
                       high = "steelblue", midpoint = 0.95,
                       limits = c(0.80, 1.00),
                       labels = scales::percent_format()) +
  facet_wrap(~re_group, scales = "free_y", ncol = 1) +
  labs(
    title    = "HAZ Coverage Heat-Map: Sample Size × RE Scenarios (Groups A–D)",
    subtitle = paste0("Term: ", heatmap_term,
                      "; nominal 95%; blue = adequate; red = under-coverage; ",
                      "reference = Malawi"),
    x = "Sample-Size Scenario", y = "RE Scenario", fill = "Coverage"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8))
```

---

# 14. Power Analysis

```{r chunk-14-power-design}
# ── Design rationale ──────────────────────────────────────────────────────────
# Power = P(reject H₀ | H₁ true) = proportion of converged reps in which
# the two-sided Wald test at α = 0.05 correctly rejects the null.
#
# In this simulation every replication uses the TRUE parameter values as the
# data-generating mechanism, so EVERY rejection is a true positive.
# Power is therefore estimated directly as:
#
#   power(term, scenario) = mean(p.value < 0.05 | converged)
#
# Three complementary views are provided:
#
#   14a. Power table — by term × sample-size scenario
#        Answers: at the real study N = 736, which terms are reliably detected?
#
#   14b. Power curves — empirical (points) + logistic-regression fit (line)
#        as a function of N for each fixed-effect term.
#        The logistic fit uses the binary reject/not-reject outcome ~ log(N),
#        giving a smooth power curve and an estimated N for 80% power.
#
#   14c. Power heat-map — term × RE scenario
#        Answers: which random-effects structures degrade power most?
#        Particularly informative for Group D (worst-case) scenarios.
#
#   14d. Power vs residual noise (Group E)
#        Answers: how much does measurement error erode detection?
#
# NOTE: α = 0.05 throughout (two-sided Wald test via lmerTest z-statistics).
# For a Bonferroni-corrected threshold (7 terms), replace 0.05 with 0.05/7.
# ─────────────────────────────────────────────────────────────────────────────
ALPHA <- 0.05
```

## 14a. Power Table by Sample Size

```{r chunk-14a-power-table}
# Compute power from the already-run sample-size scenario results
power_n_HAZ <- n_scenario_df_HAZ %>%
  filter(converged == TRUE, term %in% key_terms) %>%
  group_by(N, n_label, term) %>%
  summarise(
    n_reps        = n(),
    power         = mean(p.value < ALPHA, na.rm = TRUE),
    power_se      = sqrt(power * (1 - power) / n_reps),   # binomial SE
    .groups       = "drop"
  ) %>%
  mutate(n_label = factor(n_label,
                           levels = sapply(sample_size_scenarios, `[[`, "label")))

knitr::kable(
  power_n_HAZ %>%
    dplyr::select(n_label, term, n_reps, power, power_se) %>%
    arrange(n_label, term),
  digits  = 3,
  caption = paste0(
    "Empirical power (α = ", ALPHA, ") by fixed-effect term and sample-size scenario ",
    "(R = ", R_REPS_MAIN, " reps per cell; power_se = binomial SE)"
  )
)
```

## 14b. Power Curves with Logistic-Regression Fit

```{r chunk-14b-power-curves, fig.width = 12, fig.height = 7}
png("HAZPower Curves.png", width = 800, height = 600)
# ── Fit logistic power curve per term ────────────────────────────────────────
# Model: logit(reject) ~ log(N), fit on the raw rep-level data.
# Predicted curve is evaluated on a fine N grid for a smooth visual.
# The N at which predicted power = 0.80 is extracted and annotated.
# ─────────────────────────────────────────────────────────────────────────────

# Rep-level binary outcome for logistic regression
power_raw <- n_scenario_df_HAZ %>%
  filter(converged == TRUE, term %in% key_terms) %>%
  mutate(rejected = as.integer(p.value < ALPHA))

# Grid of N values for prediction
n_grid <- data.frame(N = seq(100, N_TARGET * 5, by = 10))

# Fit + predict per term
power_curves <- power_raw %>%
  group_by(term) %>%
  group_modify(function(df, key) {
    fit <- glm(rejected ~ log(N), data = df, family = binomial)
    pred <- predict(fit, newdata = n_grid, type = "response")

    # Estimated N for 80% power via uniroot (safe fallback if < range)
    n80 <- tryCatch({
      uniroot(function(n) predict(fit,
                                   newdata = data.frame(N = n),
                                   type    = "response") - 0.80,
              interval = c(50, N_TARGET * 10))$root
    }, error = function(e) NA_real_)

    cbind(n_grid, power_pred = pred, n80 = n80)
  }) %>%
  ungroup()

# Empirical points
power_emp <- power_n_HAZ %>%
  filter(!grepl("Registry", n_label))  # keep 3 main Ns for clarity

ggplot() +
  # Logistic fit line
  geom_line(data  = power_curves,
            aes(x = N, y = power_pred, colour = term),
            linewidth = 0.9) +
  # 80% reference
  geom_hline(yintercept = 0.80, linetype = "dashed", colour = "grey40") +
  # Empirical points with binomial CIs
  geom_point(data  = power_emp,
             aes(x = N, y = power, colour = term),
             size = 3, shape = 16) +
  geom_errorbar(data  = power_emp,
                aes(x = N,
                    ymin = pmax(0, power - 1.96 * power_se),
                    ymax = pmin(1, power + 1.96 * power_se),
                    colour = term),
                width = 30) +
  # N-for-80%-power vertical ticks per term
  geom_segment(
    data = power_curves %>%
      dplyr::select(term, n80) %>%
      distinct() %>%
      filter(!is.na(n80)),
    aes(x = n80, xend = n80, y = 0, yend = 0.80, colour = term),
    linetype = "dotted", linewidth = 0.6
  ) +
  # Real-study anchor
  geom_vline(xintercept = N_TARGET, linetype = "longdash",
             colour = "black", linewidth = 0.6) +
  annotate("text", x = N_TARGET, y = 0.05,
           label = paste0("Real study\nN = ", N_TARGET),
           hjust = -0.08, size = 3, colour = "black") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  scale_x_continuous(breaks = c(368, 736, 1000, 1472, 2500, 3680)) +
  labs(
    title    = "HAZ: Empirical Power Curves by Fixed-Effect Term",
    subtitle = paste0(
      "Lines = logistic fit [logit(reject) ~ log(N)]; points = empirical power ± 95% CI; ",
      "dashed = 80% threshold; dotted verticals = estimated N for 80% power; ",
      "α = ", ALPHA, "; R = ", R_REPS_MAIN, " reps"
    ),
    x      = "Sample size (N)",
    y      = "Power",
    colour = "Fixed-effect term"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")
dev.off()
```

## 14c. N-for-80%-Power Summary Table

```{r chunk-14c-n80-table}
n80_table <- power_curves %>%
  dplyr::select(term, n80) %>%
  distinct() %>%
  mutate(
    n80_rounded    = ceiling(n80 / 10) * 10,   # round up to next 10
    vs_real_study  = round(n80_rounded / N_TARGET, 2),
    adequately_powered_at_736 = ifelse(n80_rounded <= N_TARGET, "Yes", "No")
  ) %>%
  arrange(n80_rounded)

knitr::kable(
  n80_table,
  digits  = 1,
  col.names = c("Term", "N for 80% power (exact)",
                "N for 80% power (rounded up)",
                "As fraction of real study N",
                "Adequately powered at N = 736?"),
  caption = paste0(
    "Estimated sample size required for 80% power per fixed-effect term ",
    "(logistic power curve; α = ", ALPHA, "; baseline RE scenario)"
  )
)
```

## 14d. Power Heat-Map — Term × RE Scenario

```{r chunk-14d-power-heatmap-re, fig.width = 13, fig.height = 10}
# Compute power from the already-run RE scenario results (N = 736 throughout)
png("Power Heat-Map.png", width = 800, height = 600)
power_re_HAZ <- re_results_df_HAZ %>%
  filter(converged == TRUE, term %in% key_terms) %>%
  group_by(re_group, re_scenario, term) %>%
  summarise(
    n_reps = n(),
    power  = mean(p.value < ALPHA, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(re_scenario_clean = gsub("\n", " ", re_scenario))

ggplot(power_re_HAZ,
       aes(x = term, y = re_scenario_clean, fill = power)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = scales::percent(power, accuracy = 1)), size = 2.8) +
  scale_fill_gradient2(
    low      = "firebrick",
    mid      = "lightyellow",
    high     = "steelblue",
    midpoint = 0.80,
    limits   = c(0, 1),
    labels   = scales::percent_format()
  ) +
  facet_wrap(~re_group, scales = "free_y", ncol = 1) +
  geom_hline(yintercept = 0.80, colour = "grey60", linetype = "dashed") +
  labs(
    title    = "HAZ Power Heat-Map: Fixed-Effect Term × RE Scenario",
    subtitle = paste0(
      "N = ", N_TARGET, "; α = ", ALPHA,
      "; blue = high power; red = low power; Group D = worst-case combined; ",
      "R = ", R_REPS_MAIN, " reps"
    ),
    x    = "Fixed-effect term",
    y    = "RE Scenario",
    fill = "Power"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1, size = 8),
    axis.text.y = element_text(size = 7),
    legend.position = "right"
  )
dev.off()
```

## 14e. Power vs Residual Noise (Group E)

```{r chunk-14e-power-resid, fig.width = 11, fig.height = 6}
png("Power vs Residual Noise.png", width = 800, height = 600)
power_resid_HAZ <- resid_df_HAZ %>%
  filter(converged == TRUE, term %in% key_terms) %>%
  group_by(sigma_eps, eps_label, term) %>%
  summarise(
    n_reps   = n(),
    power    = mean(p.value < ALPHA, na.rm = TRUE),
    power_se = sqrt(power * (1 - power) / n_reps),
    .groups  = "drop"
  )

ggplot(power_resid_HAZ,
       aes(x = sigma_eps, y = power, colour = term)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = pmax(0, power - 1.96 * power_se),
                    ymax = pmin(1, power + 1.96 * power_se)),
                width = 0.03) +
  geom_hline(yintercept = 0.80, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = 0.80, linetype = "longdash",
             colour = "black", linewidth = 0.5) +
  annotate("text", x = 0.80, y = 0.03,
           label = "Baseline\n(σ_ε = 0.80)",
           hjust = -0.08, size = 3, colour = "black") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  scale_x_continuous(breaks = c(0.40, 0.60, 0.80, 1.00, 1.20, 1.60)) +
  labs(
    title    = "HAZ: Power vs Residual Noise Level (Group E)",
    subtitle = paste0(
      "N = ", N_TARGET, "; α = ", ALPHA,
      "; dashed = 80% power threshold; R = ", R_REPS_MAIN, " reps"
    ),
    x      = "Residual SD (σ_ε)",
    y      = "Power",
    colour = "Fixed-effect term"
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")
dev.off()
```

---

# 15. Reporting Summary — All Five Metrics

```{r chunk-15-summary-table}
# ── Power added to the comprehensive summary ─────────────────────────────────
# Merge power (from RE scenarios at N = 736) into the main summary.
power_re_for_merge <- power_re_HAZ %>%
  dplyr::select(re_scenario, term, power) %>%
  mutate(re_scenario_clean = gsub("\n", " ", re_scenario)) %>%
  dplyr::select(re_scenario_clean, term, power)

summary_all <- perf_re_HAZ %>%
  filter(term %in% key_terms) %>%
  mutate(
    re_scenario_clean = gsub("\n", " ", re_scenario),
    worst_case        = grepl("Worst-Case", re_scenario_clean),
    flag_bias         = abs(bias)        > 0.05,
    flag_cov          = coverage_95      < 0.93,
    flag_conv         = convergence_rate < 0.90,
    flag_power        = TRUE   # placeholder; filled after merge
  ) %>%
  left_join(power_re_for_merge,
            by = c("re_scenario_clean", "term")) %>%
  mutate(
    flag_power = power < 0.80,
    any_flag   = flag_bias | flag_cov | flag_conv | flag_power
  ) %>%
  dplyr::select(re_group, re_scenario, term, n_reps,
                bias, rmse, coverage_95, convergence_rate, power,
                flag_bias, flag_cov, flag_conv, flag_power, any_flag)

knitr::kable(
  summary_all %>%
    dplyr::select(-flag_bias, -flag_cov, -flag_conv, -flag_power) %>%
    arrange(any_flag, re_group, term),
  digits  = 3,
  caption = paste0(
    "Comprehensive HAZ performance table — Groups A–D — five metrics ",
    "(any_flag = TRUE: |bias| > 0.05, coverage < 93%, ",
    "convergence < 90%, or power < 80%; α = ", ALPHA,
    "; R = ", R_REPS_MAIN, " reps)"
  )
)
```

---

# 16. Session Info

```{r chunk-16-session}
sessionInfo()
```


```{r chunk-13i-waz-scatter, fig.width = 12, fig.height = 7}
png("HAZ trajectories.png", width = 800, height = 600)
# Run once using the baseline simulation to generate a representative dataset
set.seed(20260220)
plot_data_HAZ <- generate_HAZ_data(N = N_TARGET, dropout = 0.20) %>%
  filter(!is.na(outcome))

ggplot(plot_data_HAZ, aes(x = age_mo, y = outcome, color = country)) +
  geom_point(alpha = 0.3, size = 0.8) +
  geom_smooth(method = "loess", se = TRUE, linewidth = 1, alpha = 0.2) +
  geom_hline(yintercept =  0, linetype = "dashed", color = "red",   linewidth = 1) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "red",   linewidth = 1) +
  geom_vline(xintercept = seq(6, 60, 6), linetype = "dashed",
             color = "black", linewidth = 0.4) +
  annotate("text", x = 59, y =  0, label = "WHO expected WAZ",
           color = "black", hjust = 1.1, fontface = "italic", size = 4) +
  annotate("text", x = 59, y = -2, label = "Wasting",
           color = "black", hjust = 1.1, fontface = "italic", size = 4) +
  scale_color_manual(values = c("Uganda"       = "blue",
                                 "Malawi"       = "red",
                                 "Zimbabwe"     = "black",
                                 "South Africa" = "orange")) +
  scale_x_continuous(breaks = seq(6, 60, 6)) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Age (months)", y = "Weight-for-age z-score (WAZ)", color = "Country") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8))
dev.off()
```

```{r chunk-16a-best-scenario}
cat("\n", strrep("=", 60), "\n")
cat("Best-Performing WAZ Scenario\n")
cat(strrep("=", 60), "\n")

best_scenario_HAZ <- perf_crossed_HAZ %>%
  filter(term %in% key_terms) %>%
  group_by(N, n_label, re_group, re_scenario) %>%
  summarise(
    mean_rmse = mean(rmse,      na.rm = TRUE),
    mean_bias = mean(abs(bias), na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  arrange(mean_rmse, mean_bias) %>%
  slice(1)

cat("Sample size label :", best_scenario_HAZ$n_label,                      "\n")
cat("RE group          :", best_scenario_HAZ$re_group,                     "\n")
cat("RE scenario       :", gsub("\n", " ", best_scenario_HAZ$re_scenario), "\n")
cat("Mean RMSE         :", round(best_scenario_HAZ$mean_rmse, 4),          "\n")
cat("Mean |Bias|       :", round(best_scenario_HAZ$mean_bias, 4),          "\n")
```



```{r chunk-16b-re-params}
best_re_name_HAZ <- names(re_scenarios)[
  sapply(re_scenarios, function(sc)
    gsub("\n", " ", sc$label) == gsub("\n", " ", best_scenario_HAZ$re_scenario)
  )
]
best_re_params_HAZ <- re_scenarios[[best_re_name_HAZ]]

cat("RE scenario name  :", best_re_name_HAZ,            "\n")
cat("  sigma_b0        :", best_re_params_HAZ$sigma_b0, "\n")
cat("  sigma_b1        :", best_re_params_HAZ$sigma_b1, "\n")
cat("  rho_b01         :", best_re_params_HAZ$rho_b01,  "\n")
```



```{r}

# =============================================================================
# HAZ — BEST CROSSED SCENARIO: DATA GENERATION, LMM, GMM & EXPORT  (v2)
#
# Updated to match the improved simulation (K=1000, Groups A–D, Group E,
# coverage + convergence metrics).
#
# Selection criterion:
#   Best crossed cell = lowest mean RMSE across key_terms,
#   tie-broken by: (1) highest mean coverage_95, (2) lowest mean |bias|.
#   Cells where mean convergence_rate < 0.90 are excluded first.
#
# Prerequisites — all must exist in the environment:
#   perf_crossed_HAZ   (from chunk-12-crossed)
#   re_scenarios       (from chunk-10a-re-define)
#   key_terms          (from chunk-10c-re-performance)
#   true_values        (from chunk-6-true-values)
#   generate_HAZ_data() fit_LMM_HAZ()
# =============================================================================

library(tidyverse)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(lcmm)
library(knitr)
library(patchwork)

# pin dplyr verbs to avoid MASS / lcmm namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# ── Output directory ──────────────────────────────────────────────────────────
OUT_DIR <- "HAZ_best_scenario_outputs_v2"
if (!dir.exists(OUT_DIR)) dir.create(OUT_DIR)

# =============================================================================
# STEP 1 — Identify the best crossed scenario
#
#   Exclusion : cells with mean convergence_rate < 0.90 (unreliable estimates)
#   Primary   : lowest mean RMSE across key_terms
#   Tie-break 1: highest mean coverage_95
#   Tie-break 2: lowest mean |bias|
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("STEP 1: Identify Best Crossed Scenario\n")
cat(strrep("=", 60), "\n")

# Confirm perf_crossed_HAZ has the expected columns (v2 adds coverage_95,
# convergence_rate relative to the original script)
expected_cols <- c("N", "n_label", "re_group", "re_scenario", "term",
                   "bias", "rmse", "coverage_95", "convergence_rate")
missing_cols  <- setdiff(expected_cols, names(perf_crossed_HAZ))
if (length(missing_cols) > 0) {
  stop(
    "perf_crossed_HAZ is missing columns: ",
    paste(missing_cols, collapse = ", "),
    "\nMake sure you are using the updated simulation script (v2 / K=1000)."
  )
}

best_scenario <- perf_crossed_HAZ %>%
  filter(term %in% key_terms) %>%
  group_by(N, n_label, re_group, re_scenario) %>%
  summarise(
    mean_rmse         = mean(rmse,            na.rm = TRUE),
    mean_bias         = mean(abs(bias),       na.rm = TRUE),
    mean_coverage     = mean(coverage_95,     na.rm = TRUE),
    mean_convergence  = mean(convergence_rate, na.rm = TRUE),
    .groups           = "drop"
  ) %>%
  # Exclude poorly converging cells
  filter(mean_convergence >= 0.90) %>%
  arrange(mean_rmse, desc(mean_coverage), mean_bias) %>%
  slice(1)

if (nrow(best_scenario) == 0) {
  stop("No crossed cell survived the convergence filter (>= 90%). ",
       "Lower the threshold or inspect convergence_rate in perf_crossed_HAZ.")
}

cat("Selected scenario\n")
cat("  Sample-size label :", best_scenario$n_label,                           "\n")
cat("  RE group          :", best_scenario$re_group,                          "\n")
cat("  RE scenario       :", gsub("\n", " ", best_scenario$re_scenario),      "\n")
cat("  Mean RMSE         :", round(best_scenario$mean_rmse,        4),        "\n")
cat("  Mean |Bias|       :", round(best_scenario$mean_bias,        4),        "\n")
cat("  Mean Coverage 95% :", round(best_scenario$mean_coverage,    4),        "\n")
cat("  Mean Convergence  :", round(best_scenario$mean_convergence, 4),        "\n")

# =============================================================================
# STEP 2 — Retrieve RE parameters for the best scenario
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("STEP 2: Retrieve RE Parameters\n")
cat(strrep("=", 60), "\n")

best_re_name <- names(re_scenarios)[
  sapply(re_scenarios, function(sc)
    gsub("\n", " ", sc$label) == gsub("\n", " ", best_scenario$re_scenario)
  )
]

if (length(best_re_name) == 0) {
  stop(
    "Could not match RE scenario label '",
    gsub("\n", " ", best_scenario$re_scenario),
    "' to any entry in re_scenarios.\n",
    "Available labels:\n  ",
    paste(sapply(re_scenarios, function(s) gsub("\n", " ", s$label)),
          collapse = "\n  ")
  )
}

best_re_params <- re_scenarios[[best_re_name]]

cat("RE scenario name :", best_re_name,                  "\n")
cat("  sigma_b0       :", best_re_params$sigma_b0,       "\n")
cat("  sigma_b1       :", best_re_params$sigma_b1,       "\n")
cat("  rho_b01        :", best_re_params$rho_b01,        "\n")
cat("  RE group       :", best_re_params$group,          "\n")

# sigma_eps: the simulation crossed grid uses the default (0.80).
# If a non-default value was used for this cell, update here.
sigma_eps_final <- 0.80
cat("  sigma_eps      :", sigma_eps_final, "(default; Group E not crossed)\n")

# =============================================================================
# STEP 3 — Generate the final analysis dataset
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("STEP 3: Generate Final Dataset\n")
cat(strrep("=", 60), "\n")

set.seed(20260220)

final_HAZ_data <- generate_HAZ_data(
  N         = best_scenario$N,
  dropout   = 0.20,
  sigma_b0  = best_re_params$sigma_b0,
  sigma_b1  = best_re_params$sigma_b1,
  rho_b01   = best_re_params$rho_b01,
  sigma_eps = sigma_eps_final          # now an explicit argument in v2
) %>%
  filter(!is.na(outcome)) %>%
  mutate(
    agemos_log = log(age_mo + 1),
    agemos_inv = 1 / (age_mo + 1)
  )

# ── Sanity checks ─────────────────────────────────────────────────────────────
cat("Total observations (post-dropout) :", nrow(final_HAZ_data),                        "\n")
cat("Unique children                   :", n_distinct(final_HAZ_data$child_id),          "\n")
cat("Countries                         :", paste(sort(unique(final_HAZ_data$country)),
                                                 collapse = ", "),                        "\n")
cat("Age time-points (months)          :", paste(sort(unique(final_HAZ_data$age_mo)),
                                                 collapse = ", "),                        "\n")
cat("HAZ range                         :",
    round(min(final_HAZ_data$outcome), 2), "to",
    round(max(final_HAZ_data$outcome), 2),                                                "\n")
cat("Observations per country:\n")
print(table(final_HAZ_data$country))

# ── Export dataset ────────────────────────────────────────────────────────────
write.csv(final_HAZ_data,
          file.path(OUT_DIR, "final_HAZ_data.csv"),
          row.names = FALSE)
saveRDS(final_HAZ_data,
        file.path(OUT_DIR, "final_HAZ_data.rds"))
cat("\nDataset saved to", OUT_DIR, "(CSV + RDS)\n")

# =============================================================================
# STEP 4 — Linear Mixed Model (LMM)
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("STEP 4: Linear Mixed Model\n")
cat(strrep("=", 60), "\n")

final_model_HAZ <- fit_LMM_HAZ(final_HAZ_data)

# Convergence check
lmm_converged <- is.null(attr(final_model_HAZ, "converge_warning")) &&
                   !isSingular(final_model_HAZ)
cat("LMM convergence status:", ifelse(lmm_converged, "OK", "WARNING"), "\n")
if (!lmm_converged)
  cat("  Warning:", attr(final_model_HAZ, "converge_warning"), "\n")

cat("\n--- LMM Summary ---\n")
print(summary(final_model_HAZ))

# Fixed effects with bias, 95% CI, and Wald-test coverage flag
lmm_tidy <- broom.mixed::tidy(final_model_HAZ, effects = "fixed") %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  left_join(
    tibble(term = names(true_values), truth = unname(true_values)),
    by = "term"
  ) %>%
  mutate(
    bias        = round(estimate - truth, 4),
    ci_lo       = estimate - 1.96 * std.error,
    ci_hi       = estimate + 1.96 * std.error,
    ci_covers   = truth >= ci_lo & truth <= ci_hi,
    sig_p05     = p.value < 0.05
  )

cat("\n--- Fixed Effects vs True Values ---\n")
print(knitr::kable(
  lmm_tidy,
  digits  = 4,
  caption = paste0(
    "LMM fixed effects (estimate, SE, bias, 95% CI, coverage flag) — ",
    best_scenario$n_label, " | ",
    gsub("\n", " ", best_scenario$re_scenario)
  )
))

# Random-effects variance components
re_summary <- as.data.frame(VarCorr(final_model_HAZ))
cat("\n--- Random-Effects Variance Components ---\n")
print(re_summary)

# ── Export LMM ────────────────────────────────────────────────────────────────
write.csv(lmm_tidy,
          file.path(OUT_DIR, "lmm_fixed_effects.csv"),
          row.names = FALSE)
write.csv(re_summary,
          file.path(OUT_DIR, "lmm_random_effects.csv"),
          row.names = FALSE)
saveRDS(final_model_HAZ,
        file.path(OUT_DIR, "lmm_model_object.rds"))
cat("LMM results saved to", OUT_DIR, "\n")

# =============================================================================
# STEP 5 — Growth Mixture Model (GMM)
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("STEP 5: Growth Mixture Model\n")
cat(strrep("=", 60), "\n")

# lcmm requires a plain data.frame with integer subject ID
gmm_data <- final_HAZ_data %>%
  arrange(child_id, age_mo) %>%
  mutate(id = as.integer(factor(child_id))) %>%
  select(id,
         time       = age_mo,
         outcome,
         agemos_log,
         agemos_inv,
         country,
         sex) %>%
  as.data.frame()

cat("GMM dataset:", nrow(gmm_data), "rows |",
    n_distinct(gmm_data$id), "subjects\n")

frac_poly_formula <- outcome ~ time + agemos_log + agemos_inv

# ── 5a. 1-class reference model ───────────────────────────────────────────────
cat("\nFitting 1-class reference model...\n")
gmm_k1 <- hlme(
  fixed   = frac_poly_formula,
  random  = ~ time,
  subject = "id",
  ng      = 1,
  idiag   = FALSE,
  data    = gmm_data
)
cat("  Log-lik:", round(gmm_k1$loglik, 2), "\n")

# ── 5b. Helper: K-class GMM via gridsearch ────────────────────────────────────
fit_gmm <- function(k, ref_model, data, formula, reps = 50) {
  cat(sprintf("\nFitting %d-class GMM (gridsearch reps = %d)...\n", k, reps))
  m <- gridsearch(
    rep     = reps,
    maxiter = 20,
    minit   = ref_model,
    hlme(
      fixed    = formula,
      mixture  = ~ time + agemos_log + agemos_inv,
      random   = ~ time,
      classmb  = ~ country,
      ng       = k,
      nwg      = TRUE,
      idiag    = FALSE,
      subject  = "id",
      data     = data
    )
  )
  cat(sprintf("  Log-lik: %.2f | AIC: %.2f | BIC: %.2f\n",
              m$loglik, m$AIC, m$BIC))
  m
}

set.seed(20260220)
gmm_k2 <- fit_gmm(2, gmm_k1, gmm_data, frac_poly_formula)
gmm_k3 <- fit_gmm(3, gmm_k1, gmm_data, frac_poly_formula)
gmm_k4 <- fit_gmm(4, gmm_k1, gmm_data, frac_poly_formula)

# ── 5c. Model comparison table ────────────────────────────────────────────────
n_subj   <- n_distinct(gmm_data$id)
gmm_list <- list(K1 = gmm_k1, K2 = gmm_k2, K3 = gmm_k3, K4 = gmm_k4)

compute_entropy <- function(m) {
  pp <- m$pprob
  if (is.null(pp) || ncol(pp) <= 2) return(NA_real_)
  probs <- as.matrix(pp[, -1, drop = FALSE])
  probs[probs < 1e-10] <- 1e-10
  1 + sum(probs * log(probs), na.rm = TRUE) / (n_subj * log(ncol(probs)))
}

compute_min_class_pct <- function(m) {
  pp <- m$pprob
  if (is.null(pp) || ncol(pp) <= 2) return(100)
  cls <- apply(as.matrix(pp[, -1, drop = FALSE]), 1, which.max)
  min(table(cls)) / n_subj * 100
}

gmm_compare <- imap_dfr(gmm_list, function(m, nm) {
  tibble(
    Model         = nm,
    K             = m$ng,
    LogLik        = round(m$loglik,              2),
    AIC           = round(m$AIC,                 2),
    BIC           = round(m$BIC,                 2),
    Entropy       = round(compute_entropy(m),     3),
    Min_class_pct = round(compute_min_class_pct(m), 1)
  )
})

cat("\n--- GMM Model Comparison ---\n")
print(knitr::kable(
  gmm_compare,
  caption = "GMM fit indices — HAZ ~ time + log(time+1) + 1/(time+1)"
))

# ── 5d. Select best GMM by BIC ────────────────────────────────────────────────
best_k   <- gmm_compare %>% slice(which.min(BIC)) %>% pull(K)
best_gmm <- gmm_list[[paste0("K", best_k)]]
cat(sprintf("\nBIC-selected model: K = %d classes\n", best_k))

cat("\n--- Best GMM Summary ---\n")
print(summary(best_gmm))

cat("\n--- Posterior Classification Quality ---\n")
print(postprob(best_gmm))

# ── 5e. Assign posterior class membership ─────────────────────────────────────
post_class <- best_gmm$pprob %>%
  mutate(pred_class = apply(select(., -id), 1, which.max))

gmm_data_classed <- gmm_data %>%
  left_join(post_class %>% select(id, pred_class), by = "id")

cat("\nClass assignment counts (per subject):\n")
print(table(gmm_data_classed$pred_class[!duplicated(gmm_data_classed$id)]))

# ── 5f. Trajectory plot ───────────────────────────────────────────────────────
traj_summary <- gmm_data_classed %>%
  group_by(pred_class, time) %>%
  summarise(
    mean_HAZ = mean(outcome, na.rm = TRUE),
    se_HAZ   = sd(outcome,   na.rm = TRUE) / sqrt(n()),
    .groups  = "drop"
  ) %>%
  mutate(Class = factor(paste0("Class ", pred_class)))

p_traj <- ggplot(traj_summary,
                 aes(x = time, y = mean_HAZ, colour = Class, fill = Class)) +
  geom_ribbon(aes(ymin = mean_HAZ - 1.96 * se_HAZ,
                  ymax = mean_HAZ + 1.96 * se_HAZ),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  geom_hline(yintercept =  0, linetype = "dashed", colour = "grey50") +
  geom_hline(yintercept = -2, linetype = "dashed", colour = "firebrick",
             linewidth = 0.7) +
  annotate("text", x = 55, y = -2.1,
           label = "Stunting threshold (HAZ < -2)",
           colour = "firebrick", size = 3.2, hjust = 1) +
  scale_x_continuous(breaks = seq(0, 60, 6)) +
  labs(
    title    = sprintf("HAZ Growth Trajectories by Latent Class (K = %d)", best_k),
    subtitle = paste0(best_scenario$n_label, " | ",
                      gsub("\n", " ", best_scenario$re_scenario),
                      " | σ_ε = ", sigma_eps_final),
    x        = "Age (months)",
    y        = "Mean HAZ ± 95% CI",
    colour   = "Latent class",
    fill     = "Latent class"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "bottom")

# Posterior probability histogram
if (best_k > 1) {
  pp_long <- best_gmm$pprob %>%
    select(-id) %>%
    pivot_longer(everything(),
                 names_to  = "Class",
                 values_to = "PostProb")

  p_pp <- ggplot(pp_long, aes(x = PostProb, fill = Class)) +
    geom_histogram(binwidth = 0.05, colour = "white",
                   alpha = 0.75, position = "identity") +
    labs(title = "Posterior Class Membership Probabilities",
         x = "Posterior probability", y = "Count", fill = "Class") +
    theme_bw(base_size = 13) +
    theme(legend.position = "bottom")

  gmm_plot <- p_traj / p_pp + plot_annotation(tag_levels = "A")
} else {
  gmm_plot <- p_traj
}

print(gmm_plot)

# ── 5g. Class × country profile ───────────────────────────────────────────────
class_profile <- gmm_data_classed %>%
  filter(!duplicated(id)) %>%
  count(pred_class, country) %>%
  group_by(pred_class) %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  ungroup()

cat("\n--- Class Profile by Country ---\n")
print(knitr::kable(
  class_profile,
  col.names = c("Class", "Country", "N", "%"),
  caption   = "Class membership by country"
))

# ── 5h. Export GMM ────────────────────────────────────────────────────────────
write.csv(gmm_compare,
          file.path(OUT_DIR, "gmm_model_comparison.csv"),
          row.names = FALSE)
write.csv(gmm_data_classed,
          file.path(OUT_DIR, "gmm_data_with_classes.csv"),
          row.names = FALSE)
write.csv(class_profile,
          file.path(OUT_DIR, "gmm_class_country_profile.csv"),
          row.names = FALSE)
saveRDS(best_gmm,
        file.path(OUT_DIR, "gmm_best_model_object.rds"))
ggsave(file.path(OUT_DIR, "gmm_trajectory_plot.png"),
       plot = gmm_plot, width = 11, height = 8, dpi = 300)
cat("GMM results saved to", OUT_DIR, "\n")

# =============================================================================
# SUMMARY
# =============================================================================
cat("\n", strrep("=", 60), "\n")
cat("ANALYSIS COMPLETE\n")
cat(strrep("=", 60), "\n")
cat(sprintf("Best scenario      : %s\n",   best_scenario$n_label))
cat(sprintf("RE scenario        : %s\n",   gsub("\n", " ", best_scenario$re_scenario)))
cat(sprintf("RE group           : %s\n",   best_scenario$re_group))
cat(sprintf("sigma_eps          : %.2f\n", sigma_eps_final))
cat(sprintf("N children         : %d\n",   n_distinct(final_HAZ_data$child_id)))
cat(sprintf("N observations     : %d\n",   nrow(final_HAZ_data)))
cat(sprintf("Mean RMSE          : %.4f\n", best_scenario$mean_rmse))
cat(sprintf("Mean Coverage 95%%  : %.4f\n", best_scenario$mean_coverage))
cat(sprintf("Mean Convergence   : %.4f\n", best_scenario$mean_convergence))
cat(sprintf("LMM convergence    : %s\n",
            ifelse(lmm_converged, "OK", "WARNING")))
cat(sprintf("BIC-selected GMM   : K = %d classes\n", best_k))
cat("\nFiles written to:", normalizePath(OUT_DIR), "\n")
cat("  final_HAZ_data.csv / .rds\n")
cat("  lmm_fixed_effects.csv\n")
cat("  lmm_random_effects.csv\n")
cat("  lmm_model_object.rds\n")
cat("  gmm_model_comparison.csv\n")
cat("  gmm_data_with_classes.csv\n")
cat("  gmm_class_country_profile.csv\n")
cat("  gmm_best_model_object.rds\n")
cat("  gmm_trajectory_plot.png\n")
```

```{r}
# ─────────────────────────────────────────────────────────────
# Load libraries
# ─────────────────────────────────────────────────────────────
library(dplyr)
library(tidyr)
library(ggplot2)

# ─────────────────────────────────────────────────────────────
# 1. Age sequence
# ─────────────────────────────────────────────────────────────
agemos <- seq(0.5, 60, by = 0.5)

# ─────────────────────────────────────────────────────────────
# 2. Longitudinal model coefficients
# ─────────────────────────────────────────────────────────────
coef_class1 <- c(intercept = -1.54042, time =  0.00519,
                 agemos_log =  0.77543, agemos_inv =  0.78987)

coef_class2 <- c(intercept =  1.18302, time = -0.03727,
                 agemos_log = -0.28100, agemos_inv = -1.25924)

# ─────────────────────────────────────────────────────────────
# 3. Generate predicted trajectories
# ─────────────────────────────────────────────────────────────
pred_data <- data.frame(agemos = agemos) %>%
  mutate(
    time = agemos,
    agemos_log = log(agemos + 1),
    agemos_inv = 1 / (agemos + 1)
  ) %>%
  rowwise() %>%
  mutate(
    class1 = coef_class1["intercept"] +
             coef_class1["time"] * time +
             coef_class1["agemos_log"] * agemos_log +
             coef_class1["agemos_inv"] * agemos_inv,

    class2 = coef_class2["intercept"] +
             coef_class2["time"] * time +
             coef_class2["agemos_log"] * agemos_log +
             coef_class2["agemos_inv"] * agemos_inv
  ) %>%
  ungroup()

# ─────────────────────────────────────────────────────────────
# 4. Reshape to long format
# ─────────────────────────────────────────────────────────────
pred_long <- pred_data %>%
  pivot_longer(cols = c("class1", "class2"),
               names_to = "Class", values_to = "WAZ")

# ─────────────────────────────────────────────────────────────
# 5. Label classes (based on trajectory shape)
# ─────────────────────────────────────────────────────────────
pred_long <- pred_long %>%
  mutate(
    Class = recode(Class,
      class1 = "Improving (catch-up growth)",
      class2 = "Declining (growth faltering)"
    )
  )

# ─────────────────────────────────────────────────────────────
# 6. Plot trajectories
# ─────────────────────────────────────────────────────────────
ggplot(pred_long, aes(x = agemos, y = WAZ, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    x = "Age (months)",
    y = "Predicted height-for-age z-score (HAZ)",
    color = "Growth trajectory"
  ) +
  theme_minimal(base_size = 14)
```

```{r}

# =============================================================================
# HAZ — GMM CLASS PARAMETER TABLE (corrected coefficient extraction)
# Requires: best_gmm, best_k
# =============================================================================

library(tidyverse)
library(knitr)
library(kableExtra)

select <- dplyr::select

# =============================================================================
# PART 1 — Inspect raw coefficient names in best_gmm$best
# =============================================================================

cat("All parameter names in best_gmm$best:\n")
coef_df <- data.frame(
  name  = names(best_gmm$best),
  value = unname(best_gmm$best)
)
print(coef_df)

# =============================================================================
# PART 2 — Extract class proportions from summarytable()
#   summarytable() gives %class1, %class2, ... directly — use those
# =============================================================================

st        <- summarytable(best_gmm)
pct_cols  <- grep("^%class", colnames(st), value = TRUE)   # "%class1", "%class2", ...
class_prop <- as.numeric(st[1, pct_cols]) / 100            # convert % → proportion
names(class_prop) <- seq_along(class_prop)

cat("\nClass proportions:\n")
print(class_prop)

# =============================================================================
# PART 3 — Extract intercept & time slope from best_gmm$best
#
# Run the chunk above first and look at the printed coef_df to confirm
# the exact names. Common lcmm patterns:
#   "(Intercept) class1"  "time class1"
#   "intercept class 1"   "time class 1"
#   "class1:(Intercept)"  "class1:time"
# The regex below covers all three — adjust only if needed.
# =============================================================================

coefs     <- best_gmm$best
vc        <- tryCatch(vcov(best_gmm), error = function(e) NULL)
ses       <- if (!is.null(vc)) setNames(sqrt(diag(vc)), names(coefs)) else
               setNames(rep(NA_real_, length(coefs)), names(coefs))

get_param <- function(pattern) {
  idx <- grep(pattern, names(coefs), ignore.case = TRUE, perl = TRUE)
  if (length(idx) == 0) {
    warning("Pattern not matched: ", pattern)
    return(list(est = NA_real_, se = NA_real_))
  }
  list(est = coefs[idx[1]], se = ses[idx[1]])
}

# ── Significance stars ────────────────────────────────────────────────────────
sig_stars <- function(est, se) {
  if (any(is.na(c(est, se))) || se == 0) return("")
  z <- abs(est / se)
  dplyr::case_when(z > 3.291 ~ "***", z > 2.576 ~ "**", z > 1.960 ~ "*", TRUE ~ "")
}

# ── Build per-class list ──────────────────────────────────────────────────────
params <- lapply(seq_len(best_k), function(k) {

  # Regex covers the three most common lcmm naming conventions
  int_pat  <- sprintf(
    "intercept.{0,5}class.{0,2}%d|class.{0,2}%d.{0,5}intercept|\\(Intercept\\).{0,5}class.{0,2}%d",
    k, k, k)
  time_pat <- sprintf(
    "^time.{0,5}class.{0,2}%d|class.{0,2}%d.{0,5}\\btime\\b",
    k, k)

  int_cs  <- get_param(int_pat)
  time_cs <- get_param(time_pat)

  list(
    class    = k,
    prop     = class_prop[k],          # numeric proportion, e.g. 0.499
    int_est  = int_cs$est,
    int_se   = int_cs$se,
    time_est = time_cs$est,
    time_se  = time_cs$se
  )
})

# =============================================================================
# PART 4 — Trajectory labels & cell formatting
# =============================================================================

label_traj <- function(slope, thr = 0.005) {
  dplyr::case_when(
    is.na(slope)       ~ "Unknown",
    abs(slope) <  thr  ~ "Stable",
    slope       < -thr ~ "Declining",
    TRUE               ~ "Rising"
  )
}

fmt_est <- function(est, se) {
  if (is.na(est)) return("—")
  sprintf("%.3f%s", est, sig_stars(est, se))
}
fmt_se  <- function(se)  if (is.na(se))  "—" else sprintf("(%.3f)", se)
fmt_pct <- function(p)   sprintf("%.1f%%", as.numeric(p) * 100)

for (k in seq_along(params))
  params[[k]]$traj_label <- label_traj(params[[k]]$time_est)

# =============================================================================
# PART 5 — Assemble wide table
# =============================================================================

col_heads <- sapply(params, function(p)
  sprintf("**Class %d**<br>(%.1f%%)<br>*%s*",
          p$class, as.numeric(p$prop) * 100, p$traj_label)
)

tbl <- as.data.frame(
  rbind(
    sapply(params, function(p) fmt_est(p$int_est,  p$int_se)),
    sapply(params, function(p) fmt_se(p$int_se)),
    sapply(params, function(p) fmt_est(p$time_est, p$time_se)),
    sapply(params, function(p) fmt_se(p$time_se)),
    sapply(params, function(p) fmt_pct(p$prop))
  ),
  stringsAsFactors = FALSE
)
colnames(tbl) <- col_heads
tbl <- tbl %>% mutate(` ` = c("Estimate", "SE",
                               "Estimate", "SE",
                               "Proportion"),
                      .before = 1)

# =============================================================================
# PART 6 — Render table
# =============================================================================

gmm_param_table <- knitr::kable(
    tbl,
    align   = c("l", rep("c", best_k)),
    escape  = FALSE,
    caption = sprintf("GMM class-specific parameters — HAZ (K = %d)", best_k)
  ) %>%
  kable_classic(full_width = FALSE, html_font = "serif") %>%
  pack_rows("Intercept (baseline z-score)",          1, 2,
            bold = FALSE, italic = TRUE) %>%
  pack_rows("Time slope (z-score change per month)", 3, 4,
            bold = FALSE, italic = TRUE) %>%
  pack_rows("Class membership",                      5, 5,
            bold = FALSE, italic = TRUE) %>%
  footnote(
    general           = "*** p < 0.001  ** p < 0.01  * p < 0.05 (Wald z-test)",
    general_title     = "",
    footnote_as_chunk = TRUE
  )

print(gmm_param_table)

# =============================================================================
# PART 7 — Diagnostic: print raw values to verify extraction worked
# =============================================================================

cat("\n--- Raw extracted values ---\n")
imap(params, function(p, i) {
  cat(sprintf(
    "Class %d | prop=%.3f | intercept=%.4f (SE=%.4f) | time=%.4f (SE=%.4f) | label=%s\n",
    p$class, as.numeric(p$prop),
    p$int_est,  p$int_se,
    p$time_est, p$time_se,
    p$traj_label
  ))
})

# ── CSV export ────────────────────────────────────────────────────────────────
params_df <- imap_dfr(params, ~tibble(
  class         = .x$class,
  traj_label    = .x$traj_label,
  proportion    = round(as.numeric(.x$prop), 4),
  int_estimate  = round(.x$int_est,  4),
  int_se        = round(.x$int_se,   4),
  time_estimate = round(.x$time_est, 4),
  time_se       = round(.x$time_se,  4)
))

OUT_DIR <- "HAZ_best_scenario_outputs_v2"
if (!dir.exists(OUT_DIR)) dir.create(OUT_DIR)
write.csv(params_df,
          file.path(OUT_DIR, "gmm_class_parameters.csv"),
          row.names = FALSE)
cat("\nSaved to", file.path(OUT_DIR, "gmm_class_parameters.csv"), "\n")
```


```{r}
# ── Extract ONLY class probability columns ─────────────────────
post_probs <- gmm_k4$pprob %>%
  dplyr::select(starts_with("prob")) %>%   # <-- safer than -id
  as.matrix()

# Check rows sum to 1 (debug step)
row_sums <- rowSums(post_probs)
print(summary(row_sums))  # should be ~1

# Avoid log(0)
post_probs[post_probs == 0] <- 1e-10

# Dimensions
N <- nrow(post_probs)
K <- ncol(post_probs)

# ── Correct entropy (0–1 scale) ────────────────────────────────
entropy_HAZ <- 1 + sum(post_probs * log(post_probs)) / (N * log(K))

cat("\n=== Entropy (scaled 0–1) ===\n")
print(round(entropy_HAZ, 3))
```


